Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-22 RMS/submission.tex
proof to condition (ii)

\newcommand{\mteTask}[2]{\texttt{mteTask(}#1\texttt{,}#2\texttt{)}}
\newcommand{\deltaTask}[3]{\texttt{deltaTask(}#1\texttt{,}#2\texttt{,}#3\texttt{)}}
\newcommand{\mteIS}[1]{\texttt{mteIS(}#1\texttt{)}}
\newcommand{\deltaIS}[2]{\texttt{deltaIS(}#1\texttt{,}#2\texttt{)}}
\newcommand{\IntSrc}[3]{\texttt{<}#1\texttt{:IntSrc|val:}#2\texttt{,cycle:}#3\texttt{>}}

Several auxiliary lemmas are proved:
\begin{lemma}
Given $ID$ of sort \verb|MaybeNat| and $L$ of sort \verb|TaskList|,
\verb|mteTask(|$ID$\verb|,deltaTask(|$ID$\verb|,|$L$\verb|,|$r$\verb|))|
$=$ \verb|mteTask(|$ID$\verb|,|$L$\verb|)| \verb|monus| $r$ for all
$r\le$ \verb|mteTask(|$ID$\verb|,|$L$\verb|)|.
\end{lemma}
\begin{IEEEproof}
If $ID=$ \verb|none|, the case is trivial. By definition,
\verb|mteTask(none,deltaTask(none,|$L$\verb|,|$r$\verb|))| $=$
\verb|mteTask(none,|$L$\verb|)| $=$ \verb|INF| $=$
\verb|mteTask(none,|$L$\verb|)| \verb|monus| $r$. Otherwise, $ID=$
\verb|some| $N$ with $N$ of sort \verb|Nat|. Assume that the
\verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
\verb|deltaTask(|$ID$\verb|,|$L$\verb|,|$r$\verb|)| $= L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{ID}{\deltaTask{ID}{L}{r}}
\\  
= & \mteTask{ID}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{ID}{L}~\verb|monus|~r~.
\end{IEEEeqnarray*}
\end{IEEEproof}
\begin{lemma}
Given $ISRC$ of class \verb|IntSrc| representing an reasonable state
of interrupt source in our model, \mteIS{\deltaIS{$ISRC$}{$r$}} $=$
\mteIS{$ISRC$} \verb|monus| $r$ for all $r\le$ \mteIS{$ISRC$}.
\end{lemma}
\begin{IEEEproof}
A reasonable $ISRC$ must be of the form \IntSrc{$O$}{$v$}{$T$} with
$v\le T$. Thus by definition, 
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{ISRC}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\verb|monus|~r
= \mteIS{ISRC}~\verb|monus|~r~.
\end{IEEEeqnarray*}
\end{IEEEproof}

Now we can present the detailed proof of Theorem~\ref{t:main}:
\noindent
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i-iv) hold.
\begin{enumerate}
\item[(i)] We must prove \verb|mte(delta(|$s$\verb|,|$r$\verb|))| $=$
  \verb|mte(|$s$\verb|)| \verb|monus| $r$ for all $r\le$
  \verb|mte(|$s$\verb|)|, with $s$ being a system state of sort
  \verb|System|.
\end{enumerate}

\end{IEEEproof}

>>>>

\newcommand{\mteTask}[2]{\texttt{mteTask(}#1\texttt{,}#2\texttt{)}}
\newcommand{\deltaTask}[3]{\texttt{deltaTask(}#1\texttt{,}#2\texttt{,}#3\texttt{)}}
\newcommand{\mteIS}[1]{\texttt{mteIS(}#1\texttt{)}}
\newcommand{\deltaIS}[2]{\texttt{deltaIS(}#1\texttt{,}#2\texttt{)}}
\newcommand{\IntSrc}[3]{\texttt{<}#1\texttt{:IntSrc|val:}#2\texttt{,cycle:}#3\texttt{>}}
\newcommand{\mteIr}[1]{\texttt{mteIr(}#1\texttt{)}}
\newcommand{\mteS}[1]{\texttt{mte(}#1\texttt{)}}
\newcommand{\deltaS}[2]{\texttt{delta(}#1\texttt{,}#2\texttt{)}}

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $ID$ of sort \verb|MaybeNat| and $L$ of sort \verb|TaskList|,
\verb|mteTask(|$ID$\verb|,deltaTask(|$ID$\verb|,|$L$\verb|,|$r$\verb|))|
$=$ \verb|mteTask(|$ID$\verb|,|$L$\verb|)| \verb|monus| $r$ for all
$r\le$ \verb|mteTask(|$ID$\verb|,|$L$\verb|)|.
\end{lemma}
\begin{IEEEproof}
If $ID=$ \verb|none|, the case is trivial. By definition,
\verb|mteTask(none,deltaTask(none,|$L$\verb|,|$r$\verb|))| $=$
\verb|mteTask(none,|$L$\verb|)| $=$ \verb|INF| $=$
\verb|mteTask(none,|$L$\verb|)| \verb|monus| $r$. Otherwise, $ID=$
\verb|some| $N$ with $N$ of sort \verb|Nat|. Assume that the
\verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
\verb|deltaTask(|$ID$\verb|,|$L$\verb|,|$r$\verb|)| $= L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{ID}{\deltaTask{ID}{L}{r}}
\\  
= & \mteTask{ID}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{ID}{L}~\verb|monus|~r~.
\end{IEEEeqnarray*}
\end{IEEEproof}
\begin{lemma}
\label{l:auxis}
Given $ISRC$ of class \verb|IntSrc| representing an reasonable state
of interrupt source in our model, \mteIS{\deltaIS{$ISRC$}{$r$}} $=$
\mteIS{$ISRC$} \verb|monus| $r$ for all $r\le$ \mteIS{$ISRC$}.
\end{lemma}
\begin{IEEEproof}
A reasonable $ISRC$ must be of the form \IntSrc{$O$}{$v$}{$T$} with
$v\le T$. Thus by definition, 
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{ISRC}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\verb|monus|~r
= \mteIS{ISRC}~\verb|monus|~r~.
\end{IEEEeqnarray*}
\end{IEEEproof}
\begin{lemma}
\label{l:auxhw}
Given $HW$ of sort \verb|Hardware|, for all $r\le$ \mteIr{$HW$},
\mteIr{$HW$} $=$ \mteIr{$HW$} \verb|monus| $r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $HW$.
\end{IEEEproof}

Now we can present the detailed proof of Theorem~\ref{t:main}:
\noindent
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i-iv) hold.

Condition~(i). We must prove \verb|mte(delta(|$s$\verb|,|$r$\verb|))|
$=$ \verb|mte(|$s$\verb|)| \verb|monus| $r$ for all $r\le$
\verb|mte(|$s$\verb|)|, with $s$ being a system state of sort
\verb|System|. Let $s$ be of the form
$(\mathit{L}~\mathit{T}~\mathit{STS}~\mathit{HW}~\mathit{ISRC})$ and
$\mathit{ID}=\verb|(|\mathit{HW}\verb|).getPc|$. We only consider the
case $\mathit{ID}$\verb|::MaybeNat| in detail, while the other case
$\mathit{ID}$\verb|::Oid| is similar. By definitions of \verb|mte| and
\verb|delta|,
\begin{IEEEeqnarray*}{Cl}
  & \mteS{\deltaS{s}{r}}
\\  
= & \mteS{\deltaTask{ID}{L}{r}
\\
  & \;\;\;\;\;\;\;\;\;T~STS~HW~\deltaIS{ISRC}{r}}
\\
= & \verb|minimum(|\mteTask{ID}{\deltaTask{ID}{L}{r}}\verb|,|
\\
  & \verb|        |\mteIS{\deltaIS{ISRC}{r}}\verb|,|
\\
  & \verb|        |\mteIr{HW}\verb|)|~.
\end{IEEEeqnarray*}
By Lemmas~\ref{l:auxtask},~\ref{l:auxis} and~\ref{l:auxhw}, it can be
further reduced:
\begin{IEEEeqnarray*}{Cl}
  & \mteS{\deltaS{s}{r}}
\\  
= & \verb|minimum(|\mteTask{ID}{L}~\verb|monus|~r~\verb|,|
\\
  & \verb|        |\mteIS{ISRC}~\verb|monus|~r~\verb|,|
\\
  & \verb|        |\mteIr{HW}~\verb|monus|~r~\verb|)|
\\
= & \verb|minimum(|\mteTask{ID}{L}\verb|,|
\\
  & \verb|        |\mteIS{ISRC}\verb|,|
\\
  & \verb|        |\mteIr{HW}\verb|)|~\verb|monus|~r
\\
= & \mteS{s}~\verb|monus|~r~.
\end{IEEEeqnarray*}

Condition~(ii) follows from the fact that $r+0=r$ and
$r$~\verb|monus|~$0=r$ for all $r$.
\end{IEEEproof}
